
# Load libs
```{r message=FALSE, warning=FALSE}
library(Matrix, quietly = T)
library(dplyr, quietly = T)
library(Seurat, quietly = T)
library(SeuratDisk, quietly = T)

library(CellChat, quietly = T)
library(future, quietly = T)
library(cluster, quietly = T)
library(parallel, quietly = T)

library(ggplot2, quietly = T)
library(patchwork, quietly = T)
library(pals, quietly = T)
library(viridis, quietly = T)
library(data.table, quietly = T)
library(shades, quietly = T)

source("/home/dwm269/DWM_utils/sc_utils/seurat_helpers/seutils.R")
source("/home/dwm269/DWM_utils/sc_utils/seurat_helpers/seuplots.R")
source("/home/dwm269/DWM_utils/plotting_utils/scThemes.R")
# scTheme <- scThemes()
```

```{r}
meta <- read.csv("/workdir/dwm269/scCardiacOrganoid/resources/metadata.csv")
meta
```


# Load dataset
```{r}
rdata.path <- "/workdir/dwm269/scCardiacOrganoid/data/robjs/scCO_v7-4.rds"
if(!file.exists(rdata.path)){
  system("Rscript scripts/r/convert_h5ad_to_seurat.R") # run R script to convert anndata to seurat
}else(
  scco.seu <- readRDS(rdata.path)
)

# scco.seu$leiden_harmony_types <- factor(
#   scco.seu$leiden_harmony_types,
#   levels=c(
#     "iPSCs", 
#     "Primitive Streak",
#     "Definitive Endoderm",
#     "Gut Mesenchyme",
#     "Foregut Epithelium",
#     "Liver Progenitors",
#     "Cardiac Mesoderm",
#     "Fibroblasts",
#     "Smooth Muscle",
#     "Endocardial",
#     "Epicardial",
#     "Cardiomyocytes (early)",
#     "Cardiomyocytes (late)",
#     "Extraembryonic Mesoderm"          
#   )
# )

head(scco.seu)
```
# Plot to see that it loaded properly...
```{r}
DimPlot(
  scco.seu,
  group.by="cell_types_level2",
  reduction="phate_harmony_sp"
  # cols = 
)+
  # scTheme$umap+
  coord_fixed()
```

Split the object into each individual sample for CellChat analysis
```{r}
scco.seu.list <- SplitObject(
  scco.seu,
  split.by = "sample"
)
```


# CellChat Analysis
Link to the vignette this analysis is based on:
https://htmlpreview.github.io/?https://github.com/sqjin/CellChat/blob/master/tutorial/Comparison_analysis_of_multiple_datasets_with_different_cellular_compositions.html
```{r}
CellChatDB <- CellChatDB.human
showDatabaseCategory(CellChatDB)
```
Initialize CellChat objects
```{r}
scco.chat.list <- lapply(
  scco.seu.list,
  FUN = function (SEU) createCellChat(
    object = GetAssayData(
      SEU,
      # assay = "RNA",
      slot = "data"
    ),
    meta = data.frame(
      cell_types = SEU$cell_types,
      row.names = Cells(SEU)
    ),
    group.by = "cell_types"
  )
)
```

Run CellChat on each object, individually
```{r}
#TODO: parallelize w/ mclapply
scco.chat.list <- lapply(
  scco.chat.list,
  FUN = function(CHAT){
    # Set the database to use (whole CellChat database)
    CHAT@DB <- CellChatDB.human
    
    # subset the expression data of signaling genes for saving computation cost
    CHAT <- subsetData(
      CHAT
      # features = Features(scco.seu)[
      #   !Features(scco.seu) %in% c("H2-Q8", "H2-T9", "H2-T18", "H2-Q9" ,"H2-L", "H2-BI" ,"H2-D" ,"H60a" ,"H2-Ea-ps")
      # ]
    )
    
    # CellChat processing (memory intensive for large datasets)
    CHAT <- identifyOverExpressedGenes(CHAT)
    CHAT <- identifyOverExpressedInteractions(CHAT)
    
    return(CHAT)
  }
)

gc()
```

```{r}
# future::plan("multiprocess", workers = 20)
# options(future.globals.maxSize = 1000 * 1024^2)

scco.chat.list <- lapply(
  scco.chat.list,
  FUN = function(CHAT){
    # project gene expression data onto PPI network (optional)
    CHAT <- projectData(CHAT, PPI.human)
    
    CHAT@idents = droplevels(
      CHAT@idents, 
      exclude = setdiff(levels(meta$labels),unique(meta$labels))
    )
    
    CHAT <- computeCommunProb(CHAT)
    
    # Filter out the cell-cell communication if there are only few number of cells in certain cell groups
    CHAT <- filterCommunication(CHAT, min.cells = 10)
    
    CHAT <- computeCommunProbPathway(CHAT)
    
    CHAT <- aggregateNet(CHAT)
    
    # Compute the network centrality scores
    CHAT <- netAnalysis_computeCentrality(
      CHAT, 
      slot.name = "netP"
    ) # the slot 'netP' means the inferred intercellular communication network of signaling pathways
    
  }
)

plan("sequential") # clears memory usage from future parallelization
gc()
```

Update cell idents prior to integration
```{r}
scco.chat.list <- lapply(
  scco.chat.list,
  FUN = function(CHAT, celltypes=levels(scco.seu$cell_types)) liftCellChat(CHAT, celltypes)
)
```
Save CellChat list
```{r}
save(
  scco.chat.list,
  file="/workdir/dwm269/scCardiacOrganoid/data/robjs/scCO_v7-2_cellchat_list.RData"
)
```

## Merge and
Merge preprocessed CellChat obects
```{r}
scco.chat <- mergeCellChat(
  scco.chat.list,
  add.names = meta$sample[meta$sample %in% unique(scco.seu$sample)],
  cell.prefix = TRUE
)
```
Save CellChat object
```{r}
save(
  scco.chat,
  file="/workdir/dwm269/scCardiacOrganoid/data/robjs/scCO_v7-4_cellchat.RData"
)
```


```{r}
sessionInfo()
```

